---
title: "Early Life Adversities and Psychotic Disorders: An Investigtion of Disease Course Severity"
author: "Oskar Löf"
date: "2026-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r Libraries, echo=FALSE}
library(tidyverse)
library(car)
library(sandwich)
library(emmeans)
library(ordinal)
library(ggpubr)
library(ggstance)
```

# Load and preprocess
```{r load and preprocess data, echo=FALSE}
# Load data
df <- readRDS("placeholder.rds")

# ELA variable column names
cols1 <- c("ment_father",
          "alc_father",
          "ment_mother",
          "alc_mother",
          "parent_ill",
          "parent_div",
          "fam_conf",
          "fam_fin_prob",
          "parent_unemp",
          "subj_ill",
          "bullied",
          "assault",
          "sex_ab")
cols2 <- c("diff_conc",
          "diff_beh")
cols <- c(cols1, cols2)

# Create master data set
ELA_data <- df %>%
  dplyr::filter(age_of_onset >= 2) %>%
  dplyr::mutate(dplyr::across(-c(person, sex, Diagnostic_category), ~ as.numeric(.x))) %>% # Make variables numeric
  dplyr::mutate(sex = factor(sex, levels = c("male", "female"))) %>% # sex as factor, male as baseline
  dplyr::mutate(dplyr::across(dplyr::all_of(cols1),
                ~ dplyr::case_when(       # Re-code variables in col1
                    #...
                  ))) %>%
  dplyr::mutate(dplyr::across(dplyr::all_of(cols2),
                ~ dplyr::case_when(       # Re-code variables in col2
                    #...
                  ))) %>%
  dplyr::mutate(dplyr::across(cols2,
                ~ dplyr::if_else(
                    #...           # If no difficulties in school at all, then no difficulties in the other school specific variables either
                  ))) %>%
 dplyr::mutate(ELA_score = rowSums(dplyr::across(c(diff_conc,
                                         diff_beh,
                                         ment_father,
                                         alc_father,
                                         ment_mother,
                                         alc_mother,
                                         parent_ill,
                                         parent_div,
                                         fam_conf,
                                         fam_fin_prob,
                                         parent_unemp,
                                         subj_ill,
                                         bullied,
                                         assault,
                                         sex_ab
                                         ), ~ as.numeric(.x)), na.rm = TRUE)
         ) %>% # Create the ELA score by summing all variables
 dplyr::mutate(ref_yob = year_of_birth - min(year_of_birth, na.rm = TRUE)) %>% # Reference year of birth as a covariate
 dplyr::mutate(house_dys = rowSums(dplyr::across(c(ment_father, 
                                      alc_father,
                                         ment_mother,
                                         alc_mother,
                                         parent_ill,
                                         parent_div,
                                         fam_conf,
                                         fam_fin_prob,
                                         parent_unemp
                                         ), ~ as.numeric(.x)), na.rm = TRUE)
         ) %>% # Create Household Dysfunciton Factor
 dplyr::mutate(abuse_ill = rowSums(dplyr::across(c(subj_ill,
                                         bullied,
                                         assault,
                                         sex_ab
                                         ), ~ as.numeric(.x)), na.rm = TRUE)
         ) %>% # Create Abuse & Illness Factor
 dplyr::mutate(beh_adv = rowSums(dplyr::across(c(diff_conc,
                                         diff_beh
                                         ), ~ as.numeric(.x)), na.rm = TRUE)
         ) %>% # Create Behavioral Adversities Factor
 dplyr::mutate(Diagnostic_category = factor(Diagnostic_category, levels = c("Other psychosis", "Psychotic MDD", "Bipolar", "Schizoaffective", "Schizophrenia"), ordered = TRUE)) %>% # Factorize diagnoses according to hierarchical model
 dplyr::mutate(lev_of_edu = factor(lev_of_edu, levels = c("primary", "secondary", "tertiary"), ordered = TRUE)) %>% # And make them ordered factors
 dplyr::mutate(ELA_score_noBA = house_dys + abuse_ill) %>% # Calculating score without behavioral adversities for comparison
 dplyr::filter(rowMeans(across(all_of(cols), is.na)) <= 0.4) # Remove cases where ID hasn't answered enough questions

# Diagnosis-specific data sets
scz <- ELA_data %>% dplyr::filter(Diagnostic_category == "Schizophrenia")
sad <- ELA_data %>% dplyr::filter(Diagnostic_category == "Schizoaffective")
bd <- ELA_data %>% dplyr::filter(Diagnostic_category == "Bipolar")
mdd <- ELA_data %>% dplyr::filter(Diagnostic_category == "Psychotic MDD")
other <- ELA_data %>% dplyr::filter(Diagnostic_category == "Other psychosis")
```

# Figure presets
```{r figure presets, echo=FALSE}
# Define base figure presets
ggplot2::theme_set(ggplot2::theme_bw() +
          ggplot2::theme(
            text = ggplot2::element_text(size = 14, family = "Times New Roman"),
            legend.title = ggplot2::element_text(face = "bold"),
            legend.text = ggplot2::element_text(size = 14, family = "Times New Roman"),
            plot.title = ggplot2::element_text(face = "bold", size = 16),
            axis.title = ggplot2::element_text(face = "bold", size = 16),
            axis.text = ggplot2::element_text(size = 15)
            )
          )
```

# Psychotic diagnosis and sex distributions (Figure 1A)
```{r diag and sex dist}
# Plot data with count data for labels
plot_data <- ELA_data %>%
  dplyr::count(Diagnostic_category, sex) %>%
  dplyr::mutate(diag_prop = n / sum(n))

# Pairwise chi-square tests
table <- table(ELA_data$Diagnostic_category, ELA_data$sex)
results <- chisq.test(table)

# post-hoc testing
diags <- list(ELA_data, scz, sad, bd, mdd, other)
res <- data.frame(diag = c("total", "scz", "sad", "bd", "mdd", "other"), chi_stat = c(0, 0, 0, 0, 0, 0), p.val = c(0, 0, 0, 0, 0, 0), padj = c(0, 0, 0, 0, 0, 0))
i <- 1
for (d in diags){
  c_table <- table(d$sex)
  c <- chisq.test(c_table)
  res[i, 2] <- c$statistic
  res[i, 3] <- c$p.value
  i <- i + 1
}
# Adjust p values
res$padj <- p.adjust(res$p.val, method = "BH")
res <- res %>% dplyr::mutate(sig = dplyr::case_when(
  padj <= 0.001 ~ "***",
  padj <= 0.01 ~ "**",
  padj <= 0.05 ~ "*", 
  TRUE ~ ""))

# Plot
ggplot2::ggplot(plot_data, ggplot2::aes(x = Diagnostic_category, y = diag_prop, fill = sex, color = sex)) +
  ggplot2::geom_bar(stat = "Identity", 
           position = ggplot2::position_dodge(0.94),
           alpha = 0.4) +
  ggplot2::scale_y_continuous(limits = c(0, 0.4)) +
  ggplot2::scale_x_discrete(limits = c( 
                              "Other psychosis", 
                              "Psychotic MDD", 
                              "Bipolar", 
                              "Schizoaffective", 
                              "Schizophrenia")
                              ) +
  ggplot2::scale_fill_viridis_d(aesthetics = c("fill", "color")) +
  ggplot2::labs(title = "SUPER-Finland Psychotic Diagnosis Distribution", 
                x = NULL, 
                y = "Proportion", 
                fill = "Sex", 
                color = "Sex") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(hjust = 1, vjust = 1, angle = 15, color = "black"))

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7.5, height = 5, units = "in", dpi = 600)
```

# Age at Participation distribution (Figure 1B)
```{r age at participation dist}
# Testing for significant sex differences
chisq.test(table(ELA_data$sex))

# Statistics for summary panel
stats <- ELA_data %>%
  dplyr::select(sex, age_Questionnaire) %>%
  dplyr::summarize(mu = mean(age_Questionnaire, na.rm = TRUE), 
            median = median(age_Questionnaire, na.rm = TRUE), 
            sd = sd(age_Questionnaire, na.rm = TRUE))

# Anonymize
plot_df <- ELA_data %>%
  dplyr::mutate(
    #...
  )

#ggplot2::ggplot_build(
ggplot2::ggplot(plot_df, ggplot2::aes(x = age_Questionnaire, fill = sex, color = sex)) +
  ggplot2::geom_histogram(position = "dodge",
                          alpha = 0.4,
                          bins = 64) +
  ggplot2::labs(title = "SUPER-Finland Age and Sex Distribution", 
                x = "Age (years)", 
                y = "Count", 
                fill = "Sex", 
                color = "Sex") + 
  ggplot2::scale_fill_viridis_d(aesthetics = c("fill", "color")) +
  ggplot2::scale_x_continuous(limits = c(17, 80))#)$data[[1]]

# Counts per bars
min(table(plot_df$age_Questionnaire))
table(plot_df$sex)

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7.5, height = 5, units = "in", dpi = 600)
```

# Hierarchical clusering of ELA variables (Figure 2)
```{r hierarchical clustering}
# Create dataframe with only ELA variables
cor_dat <- ELA_data %>% 
  dplyr::select(dplyr::all_of(cols)) %>% 
  dplyr::rename(`Diff. Concentrating` = diff_conc,
                `Diff. Behavior` = diff_beh,
                `Father Psych. Disorder` = ment_father,
                `Alcoholic Father` = alc_father,
                `Mother Psych. Disorder` = ment_mother,
                `Alcoholic Mother` = alc_mother,
                `Parents Ill` = parent_ill,
                `Parents Divorced` = parent_div,
                `Fam. Financial Problems` = fam_fin_prob,
                `Parents Unemployed` = parent_unemp,
                `Family Conflict`  = fam_conf,
                `Subject Ill` = subj_ill,
                `bullied` = bullied,
                `assault` = assault,
                `Sexual Abuse` = sex_ab) %>% 
  na.omit() # Remove NAs

# Create correlation matrix
cormat <- cor(cor_dat)
cormat[is.na(cormat)] <- 0 # Remove NAs
diag(cormat) <- 1 # Make sure diagonal = 1

# Transform into distance matrix
d <- as.dist(1 - cormat)

# Run hierarchical clustering
hc <- hclust(d, method = "ward.D2")

# Create dendrogram plot
par(family = "Times New Roman",font = 1)

#tiff("placeholder.tiff", width = 3000, height = 1750, res = 300)
plot(
  hc,
  main = "Hierarchical Clustering Dendrogram of ELA Variables",
  font.main = 2,
  cex.main = 1,
  xlab = "",
  yaxt = "n",
  sub = "",
  font.lab = 2,
  cex.lab = 1.2,
  #hang = -0.1,                 # Adjust label height
  family = "Times New Roman"
)

# Manually add axis for correlations
axis(2, at = seq(0, max(hc$height), length.out = 10), labels = rev(round(seq(min(cormat), max(cormat), length.out = 10), 2)))
#dev.off()

rect.hclust(hc, k = 3, border = 2:4) # (Rectangles added to plot afterwards outside of R)
```

# ELA variable answer distributions diagnosis-wise (Figure 3)
```{r variable answers}
# All diagnostic-specific data in separate data frames in a list
dfs <- list(
  Total = ELA_data,
  SCZ = scz,
  SAD = sad,
  BD = bd,
  `Psychotic MDD` = mdd,
  `Other Psychosis` = other
)

# Convert to long format for plotting
dfs_long <- dfs %>%
  purrr::imap_dfr(~dplyr::mutate(.x, Diagnosis = .y)) %>%      # .y is the list name (df1, df2, ...)
  tidyr::pivot_longer(
    cols = dplyr::all_of(cols),
    names_to = "variable",
    values_to = "value"
  )

# Count proportional values for each variable per diagnosis
dfs_props <- dfs_long %>%
  dplyr::group_by(Diagnosis, variable, value) %>%
  dplyr::summarise(n = n(), .groups = "drop_last") %>%
  dplyr::mutate(prop = n / sum(n)) %>%
  dplyr::ungroup()

# Define order in plot and ELA categories for faceting
df_order <- c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis")
house_dys <- c("ment_father", "alc_father", "ment_mother", "alc_mother", "parent_ill", "parent_div", "fam_conf", "fam_fin_prob", "parent_unemp")
abuse_ill <- c("subj_ill", "bullied", "assault", "sex_ab")
beh_adv <- c("diff_conc", "diff_beh")

# Build final plot df
df_plot <- dfs_props %>%
 dplyr::mutate(Diagnosis = factor(Diagnosis, levels = df_order)) %>% # Make diagnosis a factor and order it by hierarchical severity model
 dplyr::mutate(category = dplyr::case_when(
    variable %in% house_dys ~ "Household Dysfunciton",              #
    variable %in% abuse_ill ~ "Abuse & Illness",                    # Apply each variable to correct ELA category
    variable %in% beh_adv ~ "Behavioral \nAdversities"              #
  ))

# Define SCZ vs rest chi-squared test data frame
df_chi_test <- ELA_data %>%
 dplyr::mutate(scz_rest = dplyr::case_when(
    Diagnostic_category == "Schizophrenia" ~ "SCZ",
    TRUE ~ "rest"))

# Chi squared test of independence
chi_sq_test <- lapply(cols, function(v) {
  tab <- table(df_chi_test[[v]], df_chi_test$scz_rest)
  test <- chisq.test(tab)
  
  data.frame(
    variable  = v,
    n         = sum(tab),
    chi_sq    = test$statistic,
    df        = test$parameter,
    p_value   = test$p.value
  )
})
df_chi_sq_res <- do.call(rbind, chi_sq_test) %>%
  dplyr::mutate(p_adjusted = p.adjust(p_value, method = "BH")) %>% # Adjust p-values
  dplyr::arrange(p_adjusted)

# Convert all significant values to one diagnosis for plotting (they are all the same within an ELA variable)
p_vals <- data.frame(variable = df_chi_sq_res$variable, p = df_chi_sq_res$p_adjusted) %>%
  dplyr::mutate(Diagnosis = "Total")

res <- p_vals %>% dplyr::mutate(sig = dplyr::case_when(
  Diagnosis == "Total" & p <= 0.001 ~ "***",
  Diagnosis == "Total" & p <= 0.01 ~ "**",
  Diagnosis == "Total" & p <= 0.05 ~ "*", 
  TRUE ~ ""))

# Adding significance data to df_plot and change to plot friendly names
df_plot <- df_plot %>% 
  dplyr::left_join(res, by = c("variable", "Diagnosis")) %>%
  dplyr::mutate(plot_var = dplyr::case_when(
    variable == "ment_father" ~ "Father Psych. Disorder",
    variable == "alc_father" ~ "Alcoholic Father",
    variable =="ment_mother" ~ "Mother Psych. Disorder",
    variable =="alc_mother" ~ "Alcoholic Mother",
    variable =="parent_ill" ~ "Parents Ill",
    variable =="parent_div" ~ "Parents Divorced",
    variable =="fam_conf" ~ "Family Conflict",
    variable =="fam_fin_prob" ~ "Fam. Financial Prob.",
    variable =="parent_unemp" ~ "Parents Unemployed",
    variable =="subj_ill" ~ "Subject Ill",
    variable =="bullied" ~ "bullied",
    variable == "assault" ~ "assault",
    variable =="sex_ab" ~ "Sexual Abuse",
    variable =="diff_conc" ~ "Diff. Concentrating",
    variable =="diff_beh" ~ "Diff. Behavior",
    TRUE ~ NA)) %>%
  dplyr::filter(value == "Yes", category %in% c("Abuse & Illness", "Household Dysfunciton", "Behavioral \nAdversities")) %>% # Only Yes answers
  dplyr::mutate(Diagnosis = factor(Diagnosis, levels = df_order))

# Check range of proportions for all variables
result <- df_plot %>%
  dplyr::group_by(variable) %>%
  dplyr::summarise(
    min_prop = min(prop),
    max_prop = max(prop),
    proportional_range = max_prop - min_prop
  )

# Plot data, faceting by Household Dysfunction, Abuse & Illness and Behavioral Adversities
p <- df_plot %>%
  dplyr::mutate(plot_var = reorder(plot_var, -prop)) %>%
  ggplot2::ggplot(ggplot2::aes(x = plot_var, y = prop, fill = Diagnosis)) +
  ggplot2::geom_col(position = ggplot2::position_dodge()) +
  ggplot2::geom_text(ggplot2::aes(x = plot_var, y = (prop + 0.15), label = sig), 
            size = 4, inherit.aes = FALSE, family = "Times New Roman") +
  ggplot2::facet_grid(~ category, scales = "free_x", space = "free_x") +
  ggplot2::scale_fill_viridis_d(direction = -1) +
  ggplot2::labs(title = "ELA Score Variables Diagnosis-wise", x = NULL, y = "Proportion") +
  ggplot2::ylim(0,1) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12),
                 axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, color = "black"),
                 strip.text = ggplot2::element_text(size = 12),
                 legend.text = ggplot2::element_text(size = 12))
print(p)

#ggplot2::ggsave("placeholder.tiff", plot = p, width = 8, height = 6, units = "in", dpi = 600)

# Smallest N in plot
count <- df_plot %>% dplyr::select(c(value, n)) %>% dplyr::filter(value == "Yes")
min(count$n)
```

# ELA score distribution  (Figure 4A)
```{r ela score dist}
# Summary statistics of the ELA score
stats <- ELA_data %>%
  dplyr::summarise(
    m = mean(ELA_score, na.rm = TRUE),
    s = sd(ELA_score, na.rm = TRUE),
    med = median(ELA_score, na.rm = TRUE),
    .groups = "drop"
  )

# Anonymize data
temp <- ELA_data %>%
 dplyr::mutate(
   #...
   )

# Barchart of ELA distribution
ggplot2::ggplot(temp, ggplot2::aes(x = factor(ELA_score), y = ggplot2::after_stat(prop), group = 1)) +
  ggplot2::geom_bar(fill = "#440154", width = 0.8) +
  ggplot2::labs(title = "ELA Score Distribution", x = "ELA Score", y = "Proportion") +
  ggplot2::ylim(0, 0.14)

# Counts per bar
table(temp$ELA_score)

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 6, height = 5, units = "in", dpi = 600)
```

# Diagnosis-wise comparison of ELA scores (Figure 4B)
```{r diag-wise emm comparison}
# ANCOVA between diagnoses
fit <- lm(ELA_score ~ Diagnostic_category + ref_yob + sex, data = ELA_data)
sm <- summary(fit)
anova(fit)
pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE) # p-value of F test

## Produce EMMs for each diagnosis
emm <- emmeans::emmeans(fit, ~ Diagnostic_category, vcov = sandwich::vcovHC(fit, type = "HC3"))
em_df <- as.data.frame(emm)

## Pairwise comparison of EMMs in plot
pairs <- as.data.frame(pairs(emm, adjust = "BH"))

# Plot
ggplot2::ggplot(data = ELA_data, ggplot2::aes(x = Diagnostic_category, y = ELA_score), fill = "ELA") +
  ggplot2::geom_boxplot(width = 0.5, 
               outlier.shape = NA,
               fill = "white",
               color = "grey70") +
  ggplot2::stat_boxplot(geom = "errorbar", width = 0.1, color = "grey70") +
  ggpubr::geom_bracket(xmin = "Schizophrenia", xmax = "Bipolar", y.position = 16, label = "***", family = "Times New Roman", label.size = 5, color = "#440154",) +
  ggpubr::geom_bracket(xmin = "Other psychosis", xmax = "Bipolar", y.position = 17.4, label = "***", family = "Times New Roman", label.size = 5, color = "#440154",) +
  ggplot2::geom_point(data = em_df,
             aes(y = emmean, fill = "EMM"),
             shape = 18,
             color = "#440154",
             size = 5) +
  ggplot2::labs(title = "ELA and EMM Score per Psychotic Diagnosis", x = NULL, y = "Score") +
  ggplot2::ylim(0, 18) +
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank(),
        legend.title = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(hjust = 1, vjust = 1, angle = 15, color = "black"))

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 5, units = "in", dpi = 600)

# N count of boxes
count <- temp %>% dplyr::select(c(ELA_score, Diagnostic_category)) %>% na.omit()
summary(count)
```

# Age at first psychosis regression (Figure 5)
## Age of onset distribution (Figure 5A)
```{r age of onset dist}
# Summary statistics
stats <- ELA_data %>%
 dplyr::mutate(age_of_onset = round(age_of_onset, 0)) %>%
  dplyr::summarise(
    m = mean(age_of_onset, na.rm = TRUE),
    s = sd(age_of_onset, na.rm = TRUE),
    med = median(age_of_onset, na.rm = TRUE),
    .groups = "drop"
  )

# Create data frame to plot and anonymize data
plot_df <- ELA_data %>%
 dplyr::mutate(age_of_onset = round(age_of_onset, 0)) %>%
 dplyr::mutate(
    # ...
  )

# Bar chart of age distribution
#ggplot2::ggplot_build(
  ggplot2::ggplot(plot_df, ggplot2::aes(x = age_of_onset)) +
  ggplot2::geom_histogram(
                 position = "dodge",
                 color = "black",
                 fill = "#440154",
                 bins = 63) +
  ggplot2::labs(title = "Age at First Psychotic Hospitalization", x = "Age (years)", y = "Count") + 
  ggplot2::scale_x_continuous(limits = c(9, 71), breaks = seq(0, 100, by = 10))#)$data[[1]]

# Counts
min(table(plot_df$age_of_onset))

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7, height = 5, units = "in", dpi = 600)
```

## Figure 5B: ELA - age regression
```{r ela-age reg}
# Run linear model
model <- lm(age_of_onset ~ 
               ELA_score +
               sex +
               ref_yob
            , data = ELA_data)

sm <- summary(model)
pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE, log.p = TRUE)

# Assumptions:
## Linearity - residual plot
plot(fitted(model), residuals(model))

## Normality - QQ-plot
car::qqPlot(residuals(model))

## Multicollinearity - VIF
car::vif(model)

# Creating new data for plotting and fixing age and sex
newdat <- data.frame(
  ELA_score = seq(-1, 20, length.out = 200),
  sex = names(which.max(table(ELA_data$sex))),
  ref_yob = mean(ELA_data$ref_yob, na.rm = TRUE)
)

# y_hat is used for the regression line in the model
newdat$y_hat <- predict(model, newdat)

# Factorize and anonymize data
temp <- ELA_data %>%
 dplyr::mutate(
   #...
   )

# Plot
ggplot2::ggplot(temp, ggplot2::aes(x = ELA_score, y = age_of_onset)) +
  ggplot2::geom_violin(trim = FALSE, fill = "#440154", color = "black", alpha = 0.7) +
  ggplot2::geom_boxplot(width = 0.25, outlier.shape = NA) +
  ggplot2::geom_line(data = subset(newdat, ELA_score >= 0 & ELA_score <= 16), 
                         ggplot2::aes(x = ELA_score, y = y_hat), 
                         colour = "red", 
                         alpha = 0.8, 
                         linetype = "dashed", 
                         linewidth = 1) +
  ggplot2::scale_x_discrete(drop = FALSE) +
  ggplot2::labs(title = "Linear Regression Between Age of Onset and ELA Score", x = "ELA Score", y = "Age (years)") +
  ggplot2::ylim(0.1, 95) +
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank())

# Counts
table(temp$ELA_score)

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7, height = 5, units = "in", dpi = 600)
```

## Diagnosis-wise forest plot (Figure 5C) and factorized forest plot (Figure 5D)
```{r ela-age diag-wise}
# List of variables to analyze
groups <- list(
   Score = "ELA_score",
  `Abuse & Illness` = "abuse_ill",
  `Behavioral Adversities` = "beh_adv",
  `Household Dysfunciton` = "house_dys",
  `Confounding Factors` = c("sexfemale", "ref_yob")
)

# List of datasets to analyze
diags <- list(ELA_data, scz, sad, bd, mdd, other)
diag_names <- c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis")

## Multicollinearity check for factorized models
model <- lm(age_of_onset ~ 
              abuse_ill +
              beh_adv +
              house_dys +
              sex +
              ref_yob
              , data = ELA_data)
car::vif(model)

# ELA score - age of onset statistics
## Loops through each dataset and applies an array of statistics to score_row_raw
score_row_raw <- dplyr::tibble()
i <- 1
for (d in diags){
  fit_tot <- lm(age_of_onset ~ 
              ELA_score +
              sex + 
              ref_yob
              , data = d)
  
  V_tot <- sandwich::vcovHC(fit_tot, type = "HC3")
  sm <- summary(fit_tot)
  
  adj_r_tot <- sm$adj.r.squared
  f_tot <- sm$fstatistic[1]
  df1_tot <- sm$fstatistic[2]
  df2_tot <- sm$fstatistic[3]
  pf_tot <- pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
  
  beta_tot <- coef(fit_tot)
  se_tot   <- sqrt(diag(V_tot))
  tval_tot <- beta_tot / se_tot
  p_tot <- 2 * pt(abs(tval_tot), df = df.residual(fit_tot), lower.tail = FALSE)
  
  score_row <- dplyr::tibble(
    term = names(beta_tot),
    estimate = as.numeric(beta_tot),
    std.error = as.numeric(se_tot),
    p.value = as.numeric(p_tot),
    crit = qt(1 - 0.05/2, df = df.residual(fit_tot)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    adj_r2 = as.numeric(adj_r_tot),
    f_val = as.numeric(f_tot),
    f_p =  as.numeric(pf_tot),
    df1 = as.numeric(df1_tot),
    df2 = as.numeric(df2_tot),
  
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
        ),
    diag = diag_names[i]
  ) %>% dplyr::filter(term == "ELA_score") # Only extract statistics related to the ELA score's regression
  
  score_row_raw <- dplyr::bind_rows(score_row_raw, score_row)
  i <- i + 1
}

# Factorized ELA score - age of onset statistics
## Loops through each dataset and applies an array of statistics to df_plot_raw
df_plot_raw = dplyr::tibble()
i <- 1
for (d in diags){
  fit <- lm(age_of_onset ~ 
              house_dys +
              beh_adv +
              abuse_ill +
              sex + 
              ref_yob
              , data = d)
  
  V <- sandwich::vcovHC(fit, type = "HC3")
  sm <- summary(fit)
  
  adj_r <- sm$adj.r.squared
  f <- sm$fstatistic[1]
  df1_val <- sm$fstatistic[2]
  df2_val <- sm$fstatistic[3]
  pf <- pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
  
  beta <- coef(fit)
  se   <- sqrt(diag(V))
  tval <- beta / se
  pval <- 2 * pt(abs(tval), df = df.residual(fit), lower.tail = FALSE)
  
  df_plot <- dplyr::tibble(
    term = names(beta),
    estimate = as.numeric(beta),
    std.error = as.numeric(se),
    statistic = as.numeric(tval),
    p.value = as.numeric(pval),
    crit = qt(1 - 0.05/2, df = df.residual(fit)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    adj_r2 = as.numeric(adj_r),
    f_val = as.numeric(f),
    f_p =  as.numeric(pf),
    df1 = as.numeric(df1_val),
    df2 = as.numeric(df2_val),
    
    # Assign each statistic to a group
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Household Dysfunciton`  ~ "Household Dysfunciton",
        term %in% groups$`Behavioral Adversities`   ~ "Behavioral Adversities",
        term %in% groups$`Abuse & Illness` ~ "Abuse & Illness",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
      ), 
    diag = diag_names[i]
    ) %>%
    dplyr::filter(term != "(Intercept)") # remove intercept
  df_plot_raw <- dplyr::bind_rows(df_plot_raw, df_plot)
  i <- i + 1
}

# Individual padding for each significance mark in plot
x_range <- range(c(df_plot_raw$conf.low, df_plot_raw$conf.high), na.rm = TRUE)
x_pad <- 0.01 * diff(x_range)

# Finalize plot data frame by combining df_plot_raw and score_row_raw
df_plot_tot <- dplyr::bind_rows(df_plot_raw, score_row_raw) %>%
  dplyr::mutate(term = reorder(term, estimate)) %>% 
  dplyr::mutate(group = factor(group, levels = c("ELA Score", "Abuse & Illness", "Behavioral Adversities", "Household Dysfunciton", "Confounding Factors"))) %>%
  dplyr::mutate(ns_x = conf.high + x_pad) %>% # Add padding for significance markers
  dplyr::filter(group != "Confounding Factors") %>%
  dplyr::mutate(diag = factor(diag, levels = c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis"))) %>%
  dplyr::mutate(diag = forcats::fct_rev(diag)) %>%
  dplyr::mutate(p_adj = p.adjust(p.value, method = "BH")) %>% # adjust p-values
  dplyr::mutate(ns_label = dplyr::case_when(
        p_adj < 0.001 ~ "***",
        p_adj < 0.01 ~ "**",       
        p_adj < 0.05 ~ "*",       
        TRUE ~ ""))

# Plot
ggplot2::ggplot(df_plot_tot, ggplot2::aes(x = estimate, y = group, color = diag)) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dotted") +
  ggstance::geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), position = ggplot2::position_dodge(width = 1.2), height = 0.5) +
  ggplot2::geom_point(size = 2.5, position = ggplot2::position_dodge(width = 1.2)) +
  ggplot2::geom_text(
    aes(x = ns_x, label = ns_label, group = diag),
    color = "black",
    size = 5,
    hjust = -0.6,
    vjust = 0.8,
    position = ggplot2::position_dodge(width = 1.2),
    family = "Times New Roman"
  ) +
  ggplot2::scale_color_viridis_d(option = "D") +
  ggplot2::labs(title = "Factorized Regression Coefficients for Age of Onset", x = "Beta (95% CI)", y = NULL, color = "Diagnostic Group") +
  ggplot2::facet_grid(group ~ ., scales = "free_y", space = "free_y", switch = "y") +
  ggplot2::guides(color = ggplot2::guide_legend(reverse = TRUE)) +
  ggplot2::theme(axis.text.y = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(size = 12),
    strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.placement = "outside",
    strip.text.y.left = ggplot2::element_text(
      size = 11,
      margin = margin(l = 10, r = 10),
      angle = 0,
      family = "Times New Roman"
    )
  )

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 6, units = "in", dpi = 600)
```

# Hospitalization Burden regression (Figure 6)
## Hospitalization distribution (Figure 6A)
```{r hosp dist}
# Summary statistics
stats <- ELA_data %>%
  dplyr::summarise(
    m = mean(hospitalization, na.rm = TRUE),
    s = sd(hospitalization, na.rm = TRUE),
    med = median(hospitalization, na.rm = TRUE),
    .groups = "drop"
  )

# Barchart of ELA distribution
ggplot2::ggplot(ELA_data, ggplot2::aes(x = hospitalization)) +
  ggplot2::geom_histogram(ggplot2::aes(y = ggplot2::after_stat(count/sum(count))), position = "dodge",
                 color = "black",
                 fill = "#440154",
                 bins = 20
                 ) +
  ggplot2::labs(title = "Distribution of Average Hospitalization Burden", x = "Avg. Hospitalization Burden", y = "Proportion") + 
  ggplot2::scale_x_continuous(limits = c(-0.1, 1.05)) +
  ggplot2::scale_y_continuous(limits = c(0, 0.109))
  
#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7, height = 5, units = "in", dpi = 600)
```

## ELA score - hospitalization burden regression (Figure 6B)
```{r ela-hosp reg}
# Run linear model
model <- lm(hospitalization ~ 
               ELA_score +
               sex +
               ref_yob
            , data = ELA_data)

sm <- summary(model)
pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)

# Assumptions:
## Linearity - residual plot
plot(fitted(model), residuals(model))

## Normality - QQ-plot
car::qqPlot(residuals(model))

## Multicollinearity - VIF
car::vif(model)

# Generating new data for plotting and fixing age and sex
newdat <- data.frame(
  ELA_score = seq(-1, 20, length.out = 200),
  sex = names(which.max(table(ELA_data$sex))),
  ref_yob = mean(ELA_data$ref_yob, na.rm = TRUE)
)

# y_hat is used for the regression line in the plot
newdat$y_hat <- predict(model, newdat)

# Factorize and anonymize data
temp <- ELA_data %>%
 dplyr::mutate(
   # ...
 )

# Plot
ggplot2::ggplot(temp, ggplot2::aes(x = ELA_score, y = hospitalization)) +
  ggplot2::geom_violin(trim = FALSE, fill = "#440154", color = "black", alpha = 0.7) +
  ggplot2::geom_boxplot(width = 0.25, outlier.shape = NA) +
  ggplot2::geom_line(data = subset(newdat, ELA_score >= 0 & ELA_score <= 16), 
                         ggplot2::aes(x = ELA_score, y = y_hat), 
                         colour = "red", 
                         alpha = 0.8, 
                         linetype = "dashed", 
                         linewidth = 1) +
  ggplot2::scale_x_discrete(drop = FALSE) +
  ggplot2::labs(title = "Linear Regression Between Hospitalization Burden & ELA Score", x = "ELA Score", y = "Avg. Hospitalization Burden") +
  ggplot2::ylim(-0.2, 1.45) +
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank())

# Counts
table(temp$ELA_score)

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7, height = 5, units = "in", dpi = 600)
```

## Diagnosis-wise forest plot (Figure 6C) and factorized forest plot (Figure 6D)
```{r ela-hosp reg diag-wise}
# Define variables
groups <- list(
   Score = "ELA_score",
  `Abuse & Illness` = "abuse_ill",
  `Behavioral Adversities` = "beh_adv",
  `Household Dysfunciton` = "house_dys",
  `Confounding Factors` = c("sexfemale", "ref_yob")
)

# Define datasets
diags <- list(ELA_data, scz, sad, bd, mdd, other)
diag_names <- c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis")

# ELA score - hospitalization burden regression statistics per diagnosis
score_row_raw <- dplyr::tibble()
i <- 1
for (d in diags){
  fit_tot <- lm(hospitalization ~ 
              ELA_score +
              sex + 
              ref_yob
              , data = d)
  
  V_tot <- sandwich::vcovHC(fit_tot, type = "HC3")
  sm <- summary(fit_tot)
  
  adj_r_tot <- sm$adj.r.squared
  f_tot <- sm$fstatistic[1]
  df1_tot <- sm$fstatistic[2]
  df2_tot <- sm$fstatistic[3]
  pf_tot <- pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
  
  beta_tot <- coef(fit_tot)
  se_tot   <- sqrt(diag(V_tot))
  tval_tot <- beta_tot / se_tot
  p_tot <- 2 * pt(abs(tval_tot), df = df.residual(fit_tot), lower.tail = FALSE)
  
  score_row <- dplyr::tibble(
    term = names(beta_tot),
    estimate = as.numeric(beta_tot),
    std.error = as.numeric(se_tot),
    p.value = as.numeric(p_tot),
    crit = qt(1 - 0.05/2, df = df.residual(fit_tot)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    adj_r2 = as.numeric(adj_r_tot),
    f_val = as.numeric(f_tot),
    f_p =  as.numeric(pf_tot),
    df1 = as.numeric(df1_tot),
    df2 = as.numeric(df2_tot),
  
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
        ),
    diag = diag_names[i]
  ) %>% dplyr::filter(term == "ELA_score")
  
  score_row_raw <- dplyr::bind_rows(score_row_raw, score_row)
  i <- i + 1
}

# Factorized ELA score - hospitalization burden regression statistics per diagnosis
df_plot_raw = dplyr::tibble()
i <- 1

for (d in diags){
  fit <- lm(hospitalization ~ 
              house_dys +
              beh_adv +
              abuse_ill +
              sex + 
              ref_yob
              , data = d)
  
  V <- sandwich::vcovHC(fit, type = "HC3")
  sm <- summary(fit)
  
  adj_r <- sm$adj.r.squared
  f <- sm$fstatistic[1]
  df1_val <- sm$fstatistic[2]
  df2_val <- sm$fstatistic[3]
  pf <- pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
  
  beta <- coef(fit)
  se   <- sqrt(diag(V))
  tval <- beta / se
  pval <- 2 * pt(abs(tval), df = df.residual(fit), lower.tail = FALSE)
  
  df_plot <- tibble(
    term = names(beta),
    estimate = as.numeric(beta),
    std.error = as.numeric(se),
    statistic = as.numeric(tval),
    p.value = as.numeric(pval),
    crit = qt(1 - 0.05/2, df = df.residual(fit)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    adj_r2 = as.numeric(adj_r),
    f_val = as.numeric(f),
    f_p =  as.numeric(pf),
    df1 = df1_val,
    df2 = df2_val,
    
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Household Dysfunciton`  ~ "Household Dysfunciton",
        term %in% groups$`Behavioral Adversities`   ~ "Behavioral Adversities",
        term %in% groups$`Abuse & Illness` ~ "Abuse & Illness",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
      ),
    diag = diag_names[i]
    ) %>%
    dplyr::filter(term != "(Intercept)")
  df_plot_raw <- dplyr::bind_rows(df_plot_raw, df_plot)
  i <- i + 1
}

# Individual padding for each significance mark in plot
x_range <- range(c(df_plot_raw$conf.low, df_plot_raw$conf.high), na.rm = TRUE)
x_pad <- 0.01 * diff(x_range)

# Finalize plot dataframe
df_plot_tot <- dplyr::bind_rows(df_plot_raw, score_row_raw) %>% 
  dplyr::mutate(term = reorder(term, estimate)) %>% 
  dplyr::mutate(group = factor(group, levels = c("ELA Score", "Abuse & Illness", "Behavioral Adversities", "Household Dysfunciton", "Confounding Factors"))) %>%
  dplyr::mutate(ns_x = conf.high + x_pad) %>%
  dplyr::filter(group != "Confounding Factors") %>%
  dplyr::mutate(diag = factor(diag, levels = c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis"))) %>%
  dplyr::mutate(diag = forcats::fct_rev(diag)) %>%
  dplyr::mutate(p_adj = p.adjust(p.value, method = "BH")) %>%
  dplyr::mutate(ns_label = dplyr::case_when(
        p_adj < 0.001 ~ "***",
        p_adj < 0.01 ~ "**",       
        p_adj < 0.05 ~ "*",       
        TRUE ~ ""))
  
ggplot2::ggplot(df_plot_tot, ggplot2::aes(x = estimate, y = group, color = diag)) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dotted") +
  ggstance::geom_errorbarh(ggplot2::aes(xmin = conf.low, xmax = conf.high),  position = ggplot2::position_dodge(width = 1.2), height = 0.5) +
  ggplot2::geom_point(size = 2.5, position = ggplot2::position_dodge(width = 1.2)) +
  ggplot2::geom_text(
    ggplot2::aes(x = ns_x, label = ns_label, group = diag),
    color = "black",
    size = 5,
    hjust = -0.6,
    vjust = 0.8,
    position = ggplot2::position_dodge(width = 1.2),
    family = "Times New Roman"
  ) +
  ggplot2::scale_color_viridis_d(option = "D") +
  ggplot2::labs(title = "Factorized Regression Coefficients \nof Hospitalization Burden", x = "Beta (95% CI)", y = NULL, color = "Diagnostic Group") +
  ggplot2::facet_grid(group ~ ., scales = "free_y", space = "free_y", switch = "y") +
  ggplot2::guides(color = ggplot2::guide_legend(reverse = TRUE)) +
  ggplot2::theme(axis.text.y = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(size = 12),
    strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
    strip.placement = "outside",
    strip.text.y.left = ggplot2::element_text(
      size = 11,
      margin = margin(l = 10, r = 10),
      angle = 0,
      family = "Times New Roman"
    )
  )

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 6, units = "in", dpi = 600)
```

# Level of Education (Figure 7 & 8)
## Level of education distribution (Figure 8A)
```{r loe dist}
# Prepare plot data
loe_data <- ELA_data %>%
  dplyr::filter(!is.na(lev_of_edu), !is.na(ELA_score)) %>%
  count(lev_of_edu) %>%
  dplyr::mutate(loe_prop = n / sum(n)) %>%
  dplyr::mutate(n_count = paste0("n = ", n))

# Barchart of LoE distribution
ggplot2::ggplot(loe_data, ggplot2::aes(x = lev_of_edu, y = loe_prop, fill = lev_of_edu)) +
  ggplot2::geom_col(width = 0.4, fill = "#440154") +
  ggplot2::scale_y_continuous(limits = c(0, 0.5)) +
  ggplot2::geom_text(ggplot2::aes(label = n_count, y = loe_prop + 0.03),
            family = "Times New Roman",
            size = 3) +
  ggplot2::labs(title = "Distribution of Level of Education",
       x = NULL,
       y = "Proportion") +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 10),
        axis.title.y = ggplot2::element_text(size = 12))

# Counts
loe_data

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 4, height = 2.5, units = "in", dpi = 600)
```

## ANCOVA and pairwise ELA EMM comparison between levels (Figure 7)
```{r ela emm diag-wise}
# Data prep
plot_df <- ELA_data %>% dplyr::select(lev_of_edu, ELA_score, ref_yob, sex)

# ANCOVA
fit <- lm(ELA_score ~ lev_of_edu + ref_yob + sex, data = plot_df)
sm <- summary(fit)
pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)

F_test <- emmeans::joint_tests(fit)
F_test$p.value

## Produce ELA EMMs for each level of education
emm <- emmeans::emmeans(fit, ~ lev_of_edu, vcov = sandwich::vcovHC(fit, type = "HC3"))
em_df <- as.data.frame(emm)

## Pairwise comparison of EMMs in plot
pairs <- as.data.frame(pairs(emm, adjust = "BH"))

# Plot
temp <- plot_df %>%
  dplyr::filter(!is.na(lev_of_edu), !is.na(ELA_score))

ggplot2::ggplot(temp, ggplot2::aes(x = lev_of_edu, y = ELA_score, fill = "ELA")) + 
  ggplot2::stat_boxplot(geom = "errorbar", width = 0.05, color = "grey70") +
  ggplot2::geom_boxplot(width = 0.33, 
               outlier.shape = NA,
               fill = "white",
               color = "grey70") +
  ggplot2::geom_point(data = em_df,
             ggplot2::aes(y = emmean, fill = "EMM"),
             shape = 18,
             color = "#440154",
             size = 5) +
  ggpubr::geom_bracket(
    xmin = c("primary", "secondary", "tertiary"),
    xmax = c("secondary", "tertiary", "primary"),
    label = c("***", "***", "***"),
    color = "#440154",
    y.position = c(16, 17.5, 19),
    family = "Times New Roman",
    label.size = 5
  ) +
  ggplot2::scale_x_discrete(drop = FALSE, labels = c("0" = "tertiary")) +
  ggplot2::labs(title = "ELA & EMM Scores per Level of Education", x = "Level of Education", y = "Score") +
  ggplot2::ylim(-1, 20) +
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank(),
        legend.title = ggplot2::element_blank(),
        axis.text = ggplot2::element_text(size = 12),
        axis.title = ggplot2::element_text(size = 14))

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 6, units = "in", dpi = 600)
```

## Level of Education Distribution (Figure 8B)
```{r}
# Factorize and anonymize data
temp <- ELA_data %>%
 dplyr::mutate(
   # ...
 )

# Generate plot data
plot_data <- temp %>%
  dplyr::filter(!is.na(lev_of_edu)) %>%
  dplyr::mutate(ELA_score = as.numeric(as.character(ELA_score))) %>%
  dplyr::group_by(ELA_score, lev_of_edu) %>%
  dplyr::summarize(count = dplyr::n(), .groups = "drop") %>%
  dplyr::group_by(lev_of_edu) %>%
  dplyr::mutate(proportion = count / sum(count)) %>%
  dplyr::ungroup()

# Calculate n for each score
n_labels <- temp %>%
  dplyr::filter(!is.na(lev_of_edu)) %>%
  dplyr::mutate(ELA_score = as.numeric(as.character(ELA_score))) %>%
  dplyr::group_by(ELA_score) %>%
  dplyr::summarise(n = dplyr::n(), .groups = 'drop') %>%
  tidyr::complete(ELA_score = 0:14, fill = list(n = 0)) %>%
  dplyr::mutate(label = paste0("n=", n))

# Plot
#ggplot2::ggplot_build(
  ggplot2::ggplot(plot_data, ggplot2::aes(x = ELA_score, y = proportion, fill = lev_of_edu)) +
  ggplot2::geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  ggplot2::scale_fill_viridis_d() +
  ggplot2::scale_x_continuous(breaks = c(0:13, 14), limits = c(-0.4, 14.5), labels = c(as.character(0:13), ">14")) +
  ggplot2::scale_y_continuous(limits = c(0, 0.169)) +
  ggplot2::labs(title = "Distribution of Level of Education",
       fill = "Level of Education",
       x = "ELA Score", 
       y = "Proportion")#)$data[[1]]

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 5, units = "in", dpi = 600)
```

## Ordinal logistic regression (Figure 8C & 8D)
```{r loe olr}
# Defining variables for regressions
groups <- list(
   Score = "ELA_score",
  `Abuse & Illness` = "abuse_ill",
  `Behavioral Adversities` = "beh_adv",
  `Household Dysfunciton` = "house_dys",
  `Confounding Factors` = c("sexfemale", "ref_yob")
)

# Defining groups for regressions
diags <- list(ELA_data, scz, sad, bd, mdd, other)
diag_names <- c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis")

# ELA score statistics
score_row_raw <- dplyr::tibble()
i <- 1
for (d in diags){
  fit_tot <- ordinal::clm(lev_of_edu ~ ELA_score,
                 nominal = ~ sex + ref_yob, 
                 data = d)
  fit_null <- ordinal::clm(lev_of_edu ~ 1,
                  data = d)
  nt_tot <- ordinal::nominal_test(fit_tot) # Checking proportional odds assumption
  nt_LRT_tot <- nt_tot$LRT[2]
  nt_p_tot <- nt_tot$`Pr(>Chi)`[2]

  aov <- anova(fit_null, fit_tot, test = "LRT")
  chi_tot <- aov$LR.stat[2]
  df_tot <- aov$df[2]
  pchi_tot <- aov$`Pr(>Chisq)`[2]
  pseudo_r2_tot <- as.numeric(logLik(fit_tot)) / as.numeric(logLik(fit_null))
  
  res <- as.data.frame(coef(summary(fit_tot)))
  OR_tot <- exp(res$Estimate)
  se_tot   <- res$`Std. Error`
  z_tot <- res$`z value`
  p_tot <- res$`Pr(>|z|)`
  
  score_row <- dplyr::tibble(
    term = rownames(res),
    nt_log_lik_tot = nt_LRT_tot,
    nt_p_tot = nt_p_tot,
    estimate = as.numeric(OR_tot),
    std.error = as.numeric(se_tot),
    z_val = as.numeric(z_tot),
    p.value = as.numeric(p_tot),
    crit = qt(1 - 0.05/2, df = df.residual(fit_tot)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    chi_sq = chi_tot,
    df = df_tot,
    p_chi = pchi_tot,
    pseudo_r2 = pseudo_r2_tot,
  
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
        ),
    diag = diag_names[i]
  ) %>% dplyr::filter(term == "ELA_score")
  
  score_row_raw <- dplyr::bind_rows(score_row_raw, score_row)
  i <- i + 1
}

# Category-specific statistics 
df_plot_raw = dplyr::tibble()
i <- 1

for (d in diags){
  fit <- ordinal::clm(lev_of_edu ~ 
              house_dys +
              beh_adv +
              abuse_ill,
              nominal = ~ sex + ref_yob
              , data = d)
  
  fit_null <- ordinal::clm(lev_of_edu ~ 1, data = d)
  
  nt <- ordinal::nominal_test(fit)
  nt_LRT1 <- nt$logLik[2]
  nt_p1 <- nt$`Pr(>Chi)`[2]
  nt_LRT2 <- nt$logLik[3]
  nt_p2 <- nt$`Pr(>Chi)`[3]
  nt_LRT3 <- nt$logLik[4]
  nt_p3 <- nt$`Pr(>Chi)`[4]
  
  aov <- anova(fit_null, fit)
  chi_sq <- aov$LR.stat[2]
  df_val <- aov$df[2]
  pchi <- aov$`Pr(>Chisq)`[2]
  pseudo_r2 <- as.numeric(logLik(fit)) / as.numeric(logLik(fit_null))
  
  res <- as.data.frame(coef(summary(fit)))
  OR <- exp(res$Estimate)
  se <- res$`Std. Error`
  z <- res$`z value`
  p <- res$`Pr(>|z|)`
  
  df_plot <- dplyr::tibble(
    term = rownames(res),
    nt_log_lik_fct = paste(nt_LRT1, nt_LRT2, nt_LRT3),
    nt_p_fct = paste(nt_p1, nt_p2, nt_p3),
    estimate = as.numeric(OR),
    std.error = as.numeric(se),
    z_val = as.numeric(z),
    p.value = as.numeric(p),
    crit = qt(1 - 0.05/2, df = df.residual(fit)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    chi_sq = chi_sq,
    df = df_val,
    p_chi = pchi,
    pseudo_r2 = pseudo_r2,
    
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Household Dysfunciton`  ~ "Household Dysfunciton",
        term %in% groups$`Behavioral Adversities`   ~ "Behavioral Adversities",
        term %in% groups$`Abuse & Illness` ~ "Abuse & Illness",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
      ),
    diag = diag_names[i]
    ) %>%
    dplyr::filter(term != "(Intercept)")
  df_plot_raw <- dplyr::bind_rows(df_plot_raw, df_plot)
  i <- i + 1
}

# Individual padding for each significance mark in plot
x_range <- range(c(df_plot_raw$conf.low, df_plot_raw$conf.high), na.rm = TRUE)
x_pad <- 0.01 * diff(x_range)

# Finalize plot data frame
df_plot_tot <- dplyr::bind_rows(df_plot_raw, score_row_raw) %>% 
  dplyr::mutate(term = reorder(term, estimate)) %>% 
  dplyr::mutate(group = factor(group, levels = c("ELA Score", "Abuse & Illness", "Behavioral Adversities", "Household Dysfunciton", "Confounding Factors"))) %>%
  dplyr::mutate(ns_x = conf.high + x_pad) %>%
  dplyr::filter(group != "Confounding Factors") %>%
  dplyr::mutate(diag = factor(diag, levels = c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis"))) %>%
  dplyr::mutate(diag = forcats::fct_rev(diag)) %>%
  dplyr::mutate(p_adj = p.adjust(p.value, method = "BH")) %>%
  dplyr::mutate(ns_label = dplyr::case_when(
        p_adj < 0.001 ~ "***",
        p_adj < 0.01 ~ "**",       
        p_adj < 0.05 ~ "*",       
        TRUE ~ ""))

# Plot
ggplot2::ggplot(df_plot_tot, ggplot2::aes(x = estimate, y = group, color = diag)) +
  ggplot2::geom_vline(xintercept = 1, linetype = "dotted") +
  ggstance::geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),  position = ggplot2::position_dodge(width = 1.2), height = 0.5) +
  ggplot2::geom_point(size = 2.5, position = position_dodge(width = 1.2)) +
  ggplot2::geom_text(
    ggplot2::aes(x = ns_x, label = ns_label, group = diag),
    color = "black",
    size = 5,
    hjust = -0.6,
    vjust = 0.8,
    position = position_dodge(width = 1.2),
    family = "Times New Roman"
  ) +
  ggplot2::scale_color_viridis_d() +
  ggplot2::scale_x_continuous(limits = c(0.28, 1.15)) +
  ggplot2::labs(title = "Factorized ORs for \nLevel of Education", x = "OR (95% CI)", y = NULL, color = "Diagnostic Group") +
  ggplot2::facet_grid(group ~ ., scales = "free_y", space = "free_y", switch = "y") +
  ggplot2::guides(color = ggplot2::guide_legend(reverse = TRUE)) +
  ggplot2::theme(axis.text.y = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(size = 12),
        strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
        strip.placement = "outside",
        strip.text.y.left = ggplot2::element_text(
          size = 11,
          margin = ggplot2::margin(l = 10, r = 10),
          angle = 0,
          family = "Times New Roman"
          )
        )

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 6, units = "in", dpi = 600)
```

# PAL errors (Figure 9)
## PAL Distribution (Figure 9A)
```{r pal dist}
# Summary statistics
stats <- ELA_data %>%
  dplyr::summarise(
    m = mean(PAL_errors, na.rm = TRUE),
    s = sd(PAL_errors, na.rm = TRUE),
    med = median(PAL_errors, na.rm = TRUE),
    .groups = "drop"
  )

# Anonymize data
plot_df <- ELA_data %>%
 dplyr::mutate(
   #...
  )

# Barchart of ELA distribution
#ggplot2::ggplot_build(
  ggplot2::ggplot(plot_df, ggplot2::aes(x = PAL_errors)) +
  ggplot2::geom_histogram(ggplot2::aes(y = ggplot2::after_stat(count/sum(count))), position = "dodge",
                 color = "black",
                 fill = "#440154",
                 bins = 23
                 ) +
  ggplot2::labs(title = "Distribution of Rank-adjusted Total PAL Errors", x = "Adjusted PAL Errors", y = "Proportion") + 
  ggplot2::scale_x_continuous(limits = c(-3, 3)) +
  ggplot2::scale_y_continuous(limits = c(0, 0.11))#)$data[[1]]

# Counts
table(plot_df$PAL_errors)

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7, height = 5, units = "in", dpi = 600)
```

## ELA score - PAL error regression (Figure 8B)
```{r ela-pal reg}
# Run linear model
model <- lm(PAL_errors ~ 
               ELA_score +
               sex +
               ref_yob
            , data = ELA_data)

# Assumptions:
## Linearity - residual plot
plot(fitted(model), residuals(model))

## Normality - QQ-plot
car::qqPlot(residuals(model))

## Multicollinearity - VIF
car::vif(model)

# Retrieve model statistics
sm <- summary(model)
pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)

# Fixing age and sex for plotting
newdat <- data.frame(
  ELA_score = seq(-1, 20, length.out = 200),
  sex = names(which.max(table(ELA_data$sex))),
  ref_yob = mean(ELA_data$ref_yob, na.rm = TRUE)
)

# y_hat is used for the model line in the plot
newdat$y_hat <- predict(model, newdat)

# Factorize and anonymize data
temp <- ELA_data %>%
 dplyr::mutate(
   #...
 )

# Plot
ggplot2::ggplot(temp, ggplot2::aes(x = ELA_score, y = PAL_errors)) +
  ggplot2::geom_violin(trim = FALSE, fill = "#440154", color = "black", alpha = 0.7) +
  ggplot2::geom_boxplot(width = 0.25, outlier.shape = NA) +
  ggplot2::geom_line(data = subset(newdat, ELA_score >= 0 & ELA_score <= 16), 
                         ggplot2::aes(x = ELA_score, y = y_hat), 
                         colour = "red", 
                         alpha = 0.8, 
                         linetype = "dashed", 
                         linewidth = 1) +
  ggplot2::scale_x_discrete(drop=FALSE) +
  ggplot2::labs(title = "Linear Regression Between Adj. Total PAL Errors & ELA Score", x = "ELA Score", y = "Adjusted PAL Errors") +
  ggplot2::ylim(-3.9, 4.4) +
  ggplot2::theme(panel.grid.minor = element_blank())

# Counts
table(temp$ELA_score)

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 7, height = 5, units = "in", dpi = 600)
```

## Diagnosis-wise and factorized forest plot of linear coefficients(Figure 9C & 9D)
```{r}
# Define variables
groups <- list(
   Score = "ELA_score",
  `Abuse & Illness` = "abuse_ill",
  `Behavioral Adversities` = "beh_adv",
  `Household Dysfunciton` = "house_dys",
  `Confounding Factors` = c("sexfemale", "ref_yob")
)

# Define groups to analyze
diags <- list(ELA_data, scz, sad, bd, mdd, other)
diag_names <- c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis")

# ELA score - PAL model Statistics
score_row_raw <- dplyr::tibble()
i <- 1
for (d in diags){
  fit_tot <- lm(PAL_errors ~ 
              ELA_score +
              sex + 
              ref_yob
              , data = d)
  
  V_tot <- sandwich::vcovHC(fit_tot, type = "HC3")
  sm <- summary(fit_tot)
  
  adj_r_tot <- sm$adj.r.squared
  f_tot <- sm$fstatistic[1]
  df1_tot <- sm$fstatistic[2]
  df2_tot <- sm$fstatistic[3]
  pf_tot <- pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
  
  beta_tot <- coef(fit_tot)
  se_tot   <- sqrt(diag(V_tot))
  tval_tot <- beta_tot / se_tot
  p_tot <- 2 * pt(abs(tval_tot), df = df.residual(fit_tot), lower.tail = FALSE)
  
  score_row <- dplyr::tibble(
    term = names(beta_tot),
    estimate = as.numeric(beta_tot),
    std.error = as.numeric(se_tot),
    p.value = as.numeric(p_tot),
    crit = qt(1 - 0.05/2, df = df.residual(fit_tot)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    adj_r2 = as.numeric(adj_r_tot),
    f_val = as.numeric(f_tot),
    f_p =  as.numeric(pf_tot),
    df1 = df1_tot,
    df2 = df2_tot,
  
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
        ),
    diag = diag_names[i]
  ) %>% dplyr::filter(term == "ELA_score")
  
  score_row_raw <- dplyr::bind_rows(score_row_raw, score_row)
  i <- i + 1
}

# ELA Category - PAL model statistics 
df_plot_raw = dplyr::tibble()
i <- 1
for (d in diags){
  fit <- lm(PAL_errors ~ 
              house_dys +
              beh_adv +
              abuse_ill +
              sex + 
              ref_yob
              , data = d)
  
  V <- sandwich::vcovHC(fit, type = "HC3")
  sm <- summary(fit)
  
  adj_r <- sm$adj.r.squared
  f <- sm$fstatistic[1]
  df1_val <- sm$fstatistic[2]
  df2_val <- sm$fstatistic[3]
  pf <- pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
  
  beta <- coef(fit)
  se   <- sqrt(diag(V))
  tval <- beta / se
  pval <- 2 * pt(abs(tval), df = df.residual(fit), lower.tail = FALSE)
  
  df_plot <- dplyr::tibble(
    term = names(beta),
    estimate = as.numeric(beta),
    std.error = as.numeric(se),
    statistic = as.numeric(tval),
    p.value = as.numeric(pval),
    crit = qt(1 - 0.05/2, df = df.residual(fit)),
    conf.low  = estimate - crit * std.error,
    conf.high = estimate + crit * std.error,
    
    adj_r2 = as.numeric(adj_r),
    f_val = as.numeric(f),
    f_p =  as.numeric(pf),
    df1 = df1_val,
    df2 = df2_val,
    
    group = dplyr::case_when(
        term %in% groups$Score ~ "ELA Score",
        term %in% groups$`Household Dysfunciton`  ~ "Household Dysfunciton",
        term %in% groups$`Behavioral Adversities`   ~ "Behavioral Adversities",
        term %in% groups$`Abuse & Illness` ~ "Abuse & Illness",
        term %in% groups$`Confounding Factors` ~ "Confounding Factors",
        TRUE ~ "Other"
      ),
    diag = diag_names[i]
    ) %>%
    dplyr::filter(term != "(Intercept)")
  df_plot_raw <- dplyr::bind_rows(df_plot_raw, df_plot)
  i <- i + 1
}

# Individual padding for each significance mark in plot
x_range <- range(c(df_plot_raw$conf.low, df_plot_raw$conf.high), na.rm = TRUE)
x_pad <- 0.01 * diff(x_range)

# Finalize plot data frame
df_plot_tot <- dplyr::bind_rows(df_plot_raw, score_row_raw) %>% 
  dplyr::mutate(term = reorder(term, estimate)) %>% 
  dplyr::mutate(group = factor(group, levels = c("ELA Score", "Abuse & Illness", "Behavioral Adversities", "Household Dysfunciton", "Confounding Factors"))) %>%
  dplyr::mutate(ns_x = conf.high + x_pad) %>%
  dplyr::filter(group != "Confounding Factors") %>%
  dplyr::mutate(diag = factor(diag, levels = c("Total", "SCZ", "SAD", "BD", "Psychotic MDD", "Other Psychosis"))) %>%
  dplyr::mutate(diag = forcats::fct_rev(diag)) %>%
  dplyr::mutate(p_adj = p.adjust(p.value, method = "BH")) %>%
  dplyr::mutate(ns_label = dplyr::case_when(
        p_adj < 0.001 ~ "***",
        p_adj < 0.01 ~ "**",       
        p_adj < 0.05 ~ "*",       
        TRUE ~ ""))
  
ggplot2::ggplot(df_plot_tot, ggplot2::aes(x = estimate, y = group, color = diag)) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dotted") +
  ggstance::geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),  position = ggplot2::position_dodge(width = 1.2), height = 0.5) +
  ggplot2::geom_point(size = 2.5, position = ggplot2::position_dodge(width = 1.2)) +
  ggplot2::geom_text(
    ggplot2::aes(x = ns_x, label = ns_label, group = diag),
    color = "black",
    size = 5,
    hjust = -0.6,
    vjust = 0.8,
    position = ggplot2::position_dodge(width = 1.2),
    family = "Times New Roman"
  ) +
  ggplot2::scale_color_viridis_d() +
  ggplot2::scale_x_continuous(limits = c(-0.15, 0.45)) +
  ggplot2::labs(title = "Factorized Regression Coefficients for \nRank-adjusted Total PAL Errors", x = "Beta (95% CI)", y = NULL, color = "Diagnostic Group") +
  ggplot2::facet_grid(group ~ ., scales = "free_y", space = "free_y", switch = "y") +
  ggplot2::guides(color = ggplot2::guide_legend(reverse = TRUE)) +
  ggplot2::theme(axis.text.y = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(size = 12),
        strip.background = ggplot2::element_rect(fill = "grey90", color = NA),
        strip.placement = "outside",
        strip.text.y.left = ggplot2::element_text(
          size = 11,
          margin = margin(l = 10, r = 10),
          angle = 0,
          family = "Times New Roman"
          )
        )

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 6, units = "in", dpi = 600)
```

# SCZ-PGS Distribution (Figure 9)
```{r pgs dist}
# Summary statistics
stats <- ELA_data %>%
  dplyr::summarise(
    m = mean(scz_prs, na.rm = TRUE),
    s = sd(scz_prs, na.rm = TRUE),
    med = median(scz_prs, na.rm = TRUE),
    .groups = "drop"
  )

# Anonymize data
plot_df <- ELA_data %>%
 dplyr::mutate(
   #...
 )

# Histogram of SCZ-PGS distribution
#ggplot2::ggplot_build(
  ggplot2::ggplot(plot_df, ggplot2::aes(x = scz_prs)) +
  ggplot2::geom_histogram(ggplot2::aes(y = ggplot2::after_stat(count/sum(count))), position = "dodge",
                 color = "black",
                 fill = "#440154",
                 bins = 20
                 ) +
  ggplot2::labs(title = "Distribution of SCZ-PGS", x = "Polygenic Score", y = "Proportion") +
  ggplot2::scale_x_continuous(limits = c(-3.2, 5)) +
  ggplot2::scale_y_continuous(limits = c(0, 0.168)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12),
        axis.title = ggplot2::element_text(size = 14))#)$data[[1]]

#ggplot2::ggsave("placeholder.tiff", plot = ggplot2::last_plot(), width = 8, height = 6, units = "in", dpi = 600)
```

# PGS model comparison table
## PGS - outcome models
```{r pgs-outcome regs}
# Multicollinearity checks - VIF
model <- lm(PAL_errors ~ 
              scz_prs +
              sex +
              ref_yob
              , data = ELA_data)
car::vif(model)
sm <- summary(model)

# beta CIs and p:s
beta <- sm$coefficients[2,1]
V <- sandwich::vcovHC(model, type = "HC3")
se <- sqrt(diag(V))[2]
tval <- beta / se
pval_pal <- 2 * pt(abs(tval), df = df.residual(model), lower.tail = FALSE)

crit <- qt(1 - 0.05 / 2, df = df.residual(model))
conf.low  <- beta - crit * se
conf.high <- beta + crit * se

# Proportional Odds assumption
model <- ordinal::clm(lev_of_edu ~ scz_prs,
                 nominal = ~ sex + ref_yob, 
                 data = ELA_data)
sm <- summary(model)
exp(sm$beta)
ordinal::nominal_test(model)

# OR CIs and p:s
res <- as.data.frame(coef(summary(model)))[7,]
OR <- exp(res$Estimate)
se <- res$`Std. Error`
z <- res$`z value`
p_loe <- res$`Pr(>|z|)`

crit <- qt(1 - 0.05 / 2, df = df.residual(model))
conf.low  <- OR - crit * se
conf.high <- OR + crit * se

# Adjust p-values
#ps <- c(p_loe, pval_hos, pval_age, pval_pal)
#p.adjust(ps, method = "BH")
```

## Model comparison (Table 2)
```{r model comp}
# Prepare model data, outcomes and variables
df <- ELA_data  
ols_outcomes <- c("age_of_onset", "hospitalization", "PAL_errors")
ord_outcome  <- "lev_of_edu"
df[[ord_outcome]] <- factor(df[[ord_outcome]], ordered = TRUE)
base_vars <- "scz_prs + sex + ref_yob"
add_var <- "ELA_score"

# Linear regression model fits' statistics
ols_rows <- lapply(ols_outcomes, function(y) {

  m1 <- lm(as.formula(paste(y, "~", base_vars)),             data = df)
  m2 <- lm(as.formula(paste(y, "~", base_vars, "+", add_var)), data = df)

  r2_base <- summary(m1)$adj.r.squared   # adj. r squared for the base model
  r2_aug  <- summary(m2)$adj.r.squared   # adj. r squared for the extended model
  delta_aic <- AIC(m2) - AIC(m1)         # delta AIC
  delta_bic <- BIC(m2) - BIC(m1)         # delta BIC

  lrt   <- anova(m1, m2)                 # likelihood ratio test
  chi2  <- lrt$`F`[2] * lrt$Df[2]        # Chi squared value
  p_val <- lrt$`Pr(>F)`[2]               # LRT p-value

  data.frame(
    Outcome    = y,
    Type       = "OLS",
    R2_Base    = r2_base,
    R2_Aug     = r2_aug,
    Delta_AIC  = delta_aic,
    Delta_BIC  = delta_bic,
    LRT_chi2   = chi2,
    p          = p_val
  )
})

## r^2 for ordinal logistic regression
get_mcfadden_adj <- function(fit, fit_null) {
  k <- length(fit$beta)                         # regression coefficients only, excludes thresholds
  1 - ((fit$logLik - k) / fit_null$logLik)      # Adj. McFadden's R^2 formula
}

# OLR model fit statistics
m_null <-  ordinal::clm(as.formula(paste(ord_outcome, "~ 1")), data = df)
m1_ord <-  ordinal::clm(lev_of_edu ~ scz_prs, nominal = ~ sex + ref_yob, data = df)
m2_ord <-  ordinal::clm(lev_of_edu ~ scz_prs + ELA_score, nominal = ~ sex + ref_yob, data = df)
lrt_ord  <- anova(m1_ord, m2_ord)
ord_row <- data.frame(
  Outcome   = paste0(ord_outcome, " (ordinal)"),
  Type      = "CLM",
  R2_Base   = get_mcfadden_adj(m1_ord, m_null),
  R2_Aug    = get_mcfadden_adj(m2_ord, m_null),
  Delta_AIC = AIC(m2_ord) - AIC(m1_ord),
  Delta_BIC = BIC(m2_ord) - BIC(m1_ord),
  LRT_chi2  = lrt_ord$LR.stat[2],
  p         = lrt_ord$`Pr(>Chisq)`[2]
)

# Put together final result table
final_table <- dplyr::bind_rows(ols_rows, ord_row) %>%
  dplyr::mutate(
    R2_Base  = round(R2_Base,  5),
    R2_Aug   = round(R2_Aug,   5),
    Delta_AIC = round(Delta_AIC, 5),
    Delta_BIC = round(Delta_BIC, 5),
    LRT_chi2  = round(LRT_chi2, 5),
    adj_p    = signif(p.adjust(p, method = "BH"), 5)
  ) %>%
  dplyr::select(-Type) %>%
  dplyr::rename(
    "Outcome"          = Outcome,
    "R² Base"          = R2_Base,
    "R² Aug"           = R2_Aug,
    "d_AIC"             = Delta_AIC,
    "d_BIC"             = Delta_BIC,
    "LRT X²"           = LRT_chi2,
    "p"                = p
  )
print(final_table, row.names = FALSE)
```